---
title: Ebola Sierra Leone analysis
author: Lawal Omotola 
date: 2024-06-29
date-format: short
---

# Load packages ----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

if(!require(pacman)) install.packages("pacman")
pacman::p_load(
  tidyverse, # meta-package
  inspectdf,
  plotly,
  janitor,
  visdat,
  esquisse,
  reactable
)
```

# Load data ----

```{r}
library(readr)
ebola_sierra_leone <- read_csv("C:/Users/Hon.Olayinka/Documents/GitHub/tv_hours_pages/ebola_sierra_leone.csv")
```

# Explore data ----

```{r}
head(ebola_sierra_leone)
ncol(ebola_sierra_leone)
nrow(ebola_sierra_leone)
dim(ebola_sierra_leone)
summary(ebola_sierra_leone)


visdat::vis_dat(ebola_sierra_leone)
```

# Categorical overview ----

```{r}
cat_summary <- inspect_cat(ebola_sierra_leone)
cat_summary_plot <- show_plot(cat_summary) # function nesting

ggplotly(cat_summary_plot)
```

# Numerical overview ----

```{r}
num_summary <- inspect_num(ebola_sierra_leone)
num_summary_plot <- show_plot(num_summary)
ggplotly(num_summary_plot)
```

# Analyzing individual variables: numeric ----

```{r}
ebola_sierra_leone$age
mean(ebola_sierra_leone$age, na.rm = TRUE)
median(ebola_sierra_leone$age, na.rm = T)
age_vec <- ebola_sierra_leone$age
sd(age_vec, na.rm = T)
min(age_vec, na.rm = T)
max(age_vec, na.rm = T)
summary(age_vec, na.rm = T)
length(age_vec)
sum(age_vec, na.rm = T)

# Visualizing individual variables: numeric ----
hist(age_vec)
boxplot(age_vec)

```

# Analyzing single categorical variable: ----

```{r}
ebola_sierra_leone$district

table(ebola_sierra_leone$district)

janitor::tabyl(ebola_sierra_leone, district, sex)
```

# Analyzing single categorical variable: ----

```{r}
barplot(table(ebola_sierra_leone$district))

esquisser(ebola_sierra_leone)

ggplot(ebola_sierra_leone) +
  aes(x = district) +
  geom_bar(fill = "#112446") +
  theme_minimal()
```

# Questions ----

# When was the first case reported? (Hint: look at the date of sample)

```{r}
min(ebola_sierra_leone$date_of_sample)
```

# What was the median age of those affected?

```{r}
median(ebola_sierra_leone$age, na.rm = T)
```

# Had there been more cases in men or women?

```{r}
janitor::tabyl(ebola_sierra_leone$sex)
```

# What district had had the most reported cases?

```{r}
janitor::tabyl(ebola_sierra_leone$district)
```

# By the end of June 2014, was the outbreak growing or receding?

```{r}
ggplot(ebola_sierra_leone) +
  aes(x = date_of_onset) +
  geom_histogram(bins = 30L, fill = "#112446") +
  theme_minimal()
```
